---
title: "Creating CINeMA .cnm files from Bayesian NMA (MetaInsight / gemtc)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating CINeMA .cnm files from Bayesian NMA (MetaInsight / gemtc)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`cinemar` turns a network meta-analysis into a CINeMA v3 exchange file (`.cnm`)
that you upload directly into [CINeMA](https://cinema.med.auth.gr). It targets
the MetaInsight workflow: a **Bayesian** NMA fitted with
[gemtc](https://cran.r-project.org/package=gemtc).

## Why both frameworks appear in one file

CINeMA's confidence domains need two kinds of quantity:

* **Effect estimates and intervals** — the treatment effects, credible /
  predictive intervals, nodesplit (incoherence) and between-study SD (tau). A
  Bayesian model provides these.
* **Structural quantities** — the hat / contribution matrix, prop-direct, the
  design-by-treatment inconsistency test and pairwise heterogeneity. These are
  frequentist constructs that a Bayesian model cannot produce.

So `cinemar` always runs `netmeta` for the structural block, and — when you
supply a Bayesian fit — writes the posterior results to the Bayesian block.
On upload, CINeMA promotes the Bayesian block to primary (effects, intervals,
nodesplit, tau, league table) and keeps the frequentist block for the
structural quantities. This is the schema's *preference rule*.

## A complete MetaInsight-style example

The `smoking` dataset shipped with gemtc is MetaInsight's default example
(binary outcome, odds ratios).

```{r}
library(gemtc)
library(cinemar)

data(smoking)
ab <- smoking$data.ab   # columns: study, treatment, responders, sampleSize

# --- Bayesian NMA, exactly as MetaInsight fits it ------------------------
net   <- mtc.network(ab)
model <- mtc.model(net, likelihood = "binom", link = "logit",
                   linearModel = "random")
res   <- mtc.run(model, n.adapt = 2000, n.iter = 10000)
ranks <- rank.probability(res, preferredDirection = 1)   # for SUCRA
```

`cinemar` needs the study data in long form, plus per-arm risk of bias and
indirectness (1 = low, 2 = moderate, 3 = high) for CINeMA's domains:

```{r}
D <- data.frame(
  study = ab$study, id = ab$study, t = ab$treatment,
  r = ab$responders, n = ab$sampleSize,
  rob = 1L, indirectness = 1L
)

write_cinema_cnm(
  data     = D,
  type     = "long_binary",
  sm       = "OR",
  model    = "random",
  bayesian = res,      # gemtc mtc.result  -> Bayesian block (primary)
  ranks    = ranks,    # rank probabilities -> SUCRA
  title    = "Smoking cessation (Bayesian)",
  file     = "smoking_bayesian.cnm"
)
```

Upload `smoking_bayesian.cnm` into CINeMA: the league table, imprecision and
incoherence assessments use the posterior estimates, while the contribution
matrix (within-study bias, indirectness) comes from netmeta.

## Bringing your own estimates

If you already hold posterior summaries (MetaInsight does), skip the gemtc
extraction and pass them straight in. Comparison ids are `"A:B"` (alphabetical)
and report the effect of **A relative to B**:

```{r}
comps <- data.frame(
  comparison = c("A:B", "A:C", "A:D", "B:C", "B:D", "C:D"),
  effect   = c(...), ciLower = c(...), ciUpper = c(...),
  priLower = c(...), priUpper = c(...)
)
bblock <- cnm_bayesian_block(
  comps,
  tau = list(mean = ., ciLower = ., ciUpper = .),
  dic = list(Dbar = ., pD = ., DIC = ., dataPoints = .)
)

freq <- cnm_frequentist_block(D, type = "long_binary", sm = "OR")
write_cnm(build_cnm(freq, bayesian = bblock, title = "Smoking cessation"),
          "smoking_bayesian.cnm")
```

## Frequentist-only files

Omit `bayesian` to write a frequentist `.cnm` (equivalent to CINeMA's
*Download R script* export):

```{r}
write_cinema_cnm(D, type = "long_binary", sm = "OR", title = "Smoking (FE/RE)",
                 file = "smoking_frequentist.cnm")
```

## Orientation note

gemtc's basic parameters `d.A.B` are the effect of B relative to A; CINeMA
reports `"A:B"` as A relative to B, so `cnm_bayesian_gemtc()` negates them to
match. The orientation is verified against the frequentist league table in the
package tests (signs agree for every comparison on the smoking network). If a
future gemtc version changes this, pass `flip = TRUE`.
